---
title: 'IMT 573: PSet08 - Predictive Modeling'
author: "Iu-Yu Melody Chang"
output: html_document
---

# Instructions

Before beginning this assignment, please ensure you have access to R
and RStudio.

1. Update the PSet number and title in the `title:` field.

1. Replace the "Insert Your Name Here" text in the `author:` field
   with your own full name. 

1. Be sure to include well-documented (e.g. commented) code chucks,
   figures and clearly written text-chunk explanations as
   necessary. Any figures should be clearly labeled and appropriately
   referenced within the text.

1. When you have completed the assignment and have **checked** that
   your code both runs in the Console and knits correctly when you
   click `Knit to HTML`, rename the R Markdown file to
   `YourLastName-YourFirstName-psetX.Rmd`, where `X` is the PSet number, 
   knit the .rmd file as an html, and submit both the .rmd and the .html files 
   on Canvas.
   
1. List any collaborators in the section below.

# Collaborators
- Insert collaborator name here

# Setup

Do whatever setup you do here, such as loading libraries

```{r setup, message=FALSE}
# Load standard libraries
library("tidyverse")
```

# Question 1

## Question 1.1
Load the data and check the basic sanity.
```{r}
compas<- read.delim("compas-score-data.csv.bz2")
summary(compas)
any(is.na(compas))
```

## Question 1.2
Select only the "African-American", "Caucasian" category
```{r}
compas0 <- compas %>%
  filter(race %in% c("African-American", "Caucasian"))
```

## Question 1.3
Create a dummy variable
```{r}
compas0 <- compas0 %>%
  mutate(
    risk = cut(
      decile_score,
      breaks = c(1, 4, 10),  
      labels = c("low risk", "high risk"),
      include.lowest = TRUE,
      right = TRUE))
```

## Question 1.4
a.)The recidivism rate for low-risk individuals is 0.2264 while for high-risk individuals is 0.7195
b.)The recidivism rate for Caucasian is 0.3635 while for African-American is 0.5276
```{r}

risk_recidivism_rates <- compas0 %>%
  group_by(risk) %>%
  summarize(mean(decile_score, na.rm = TRUE)/10)
print(risk_recidivism_rates)

race_recidivism_rates <- compas0 %>%
  group_by(race) %>%
  summarize(mean(decile_score, na.rm = TRUE)/10)
print(race_recidivism_rates)

```

## Question 1.5
Accuracy=65.8% The proportion of correct predictions among all predictions is 65.8%

Precision=63.45% The proportion of true positive predictions among all positive predictions is 63.45%. In this context, it represents the accuracy of predicting recidivism among those classified as low risk.

Recall=64.52% The proportion of true positive predictions among all actual positives is 64.52%. In this context, it represents the ability to capture recidivists among all individuals who actually recidivated.
```{r}
conf_matrix <- table(Actual = compas0$two_year_recid, Predicted = compas0$risk)
print(conf_matrix)

true_positive <- conf_matrix[2, 2]  # Actual recidivist correctly classified as low risk
false_positive <- conf_matrix[1, 2]  # Non-recidivist incorrectly classified as low risk
true_negative <- conf_matrix[1, 1]  # Non-recidivist correctly classified as low risk
false_negative <- conf_matrix[2, 1]  # Actual recidivist incorrectly classified as low risk

compas_accuracy <- ((true_positive + true_negative) / sum(conf_matrix)) 
compas_accuracy
compas_precision <- ((true_positive) / (true_positive+false_positive))
compas_precision
compas_recall <- ((true_positive) / (false_negative+true_positive)) 
compas_recall

```

## Question 1.6
According to the results presented above, the actual low recidivism is predicted as high risk, while the actual high recidivism is predicted as low risk.

Therefore, I don't think this is an good model because there are too many flaws and may misunderstand the ones who are not recidivism to be risky, which is not fair.

## Question 1.7

```{r}
#"African-American"
compas1 <- compas0 %>%
  filter(race %in% c("African-American"))

conf_matrix1 <- table(Actual = compas1$two_year_recid, Predicted = compas1$risk)
print(conf_matrix1)

true_positive1 <- conf_matrix1[2, 2] 
false_positive1 <- conf_matrix1[1, 2]  
true_negative1 <- conf_matrix1[1, 1]  
false_negative1 <- conf_matrix1[2, 1]  

compas_accuracy1 <- ((true_positive1 + true_negative1) / sum(conf_matrix1)) 
FPR1<-false_positive1/(false_positive1+true_negative1)
FNR1<-false_negative1/(false_negative1+true_positive1)
compas_accuracy1
FPR1
FNR1

#"Caucasian"
compas2 <- compas0 %>%
  filter(race %in% "Caucasian")

conf_matrix2 <- table(Actual = compas2$two_year_recid, Predicted = compas2$risk)
print(conf_matrix2)

true_positive2 <- conf_matrix2[2, 2] 
false_positive2 <- conf_matrix2[1, 2]  
true_negative2 <- conf_matrix2[1, 1]  
false_negative2 <- conf_matrix2[2, 1]  

compas_accuracy2 <- ((true_positive2 + true_negative2) / sum(conf_matrix2)) 
FPR2<-false_positive2/(false_positive2+true_negative2)
FNR2<-false_negative2/(false_negative2+true_positive2)
compas_accuracy2
FPR2
FNR2

```

## Question 1.8
"African-American" Accuracy=64.91%, FPR=42.34%, FNR=28.48%

"Caucasian" Accuracy=67.19%, FPR=22.01%, FNR=49.64%

The accuracy here is similar in both African American and Caucasian, the FPR is higher in African American while the FNR is higher in Caucasian.

FPR is the proportion of actual negatives (non-events) that are incorrectly predicted as positive (events) by the model. The FPR in "African-American" is 42% that means it is often predicted incorrectly.
FNR is the proportion of actual positives (events) that are incorrectly predicted as negative (non-events) by the model. The FNR in "African-American" is 49% that means it is often predicted incorrectly.

I think it's unfair, because it is predicted incorrectly for many cases, actual low risk be predicted as high risk while actual high risk predicted as low risks. Then we may misunderstand some nice people who really wants to may changes.Also, when doing research, we do not want to misunderstand anyone. 

I think in reality, there are differences between races, so it is normal to have those differences between races.

